### 2014-06-08 latex/paper2.tex
have 6 pages

\begin{document}
\begin{center}
\textbf{P. A. M. Dirac and His Search for Beautiful Mathematics}
\end{center}

@@

Dirac began studying engineering at the University of Bristol when he
was 16. During his time at the university, he read in \textit{The Times}
about the verification of Einstein's general theory of relativity, which
had a huge influence on him and sparked his interest in physics. Eager
to find out more about Einstein's theory, Dirac was disappointed when he
realized that none of his teachers knew much about relativity either.
Fortunately for Dirac, the philosopher Charlie Broad, who had studied the
Einstein's theory, was appointed as the Professor of Philosphy at
Bristol in Dirac's final year, and gave lectures on relativity. From
these lectures and from reading Arthur Eddington's book on relativity,
Dirac was able to learn both the special and general theory of
relativity.

While Einstein's theories sparked Dirac's interest in physics, Dirac's
education in engineering helped him develop an appreciation for
approximate theories, because engineers had to constantly make
approximations. Dirac realized that Newtonian mechanics can be viewed as
an approximation to Einstein's more refined relativistic mechanics, and
perhaps Einstein's theory is also an approximation to reality. Dirac
developed the belief that the purpose of science is to develop ever more
accurate approximations, and that an exact theory might never be found.
Dirac thought that ``even approximate theories can have mathematical
beauty'' \cite[pg. 45]{strangest_man} as a result of his engineering training,
and this point of view of science as approximations has influenced
Richard Feynman and many other physicists today.

@@

Dirac was a hard-working student at Cambridge. He worked almost all day
for six days every week, reading books, journals, and reviewing his
lecture notes, only taking time off on Sundays. On Sundays, he would
take long walks that lasted the whole day, and tried not to think about
his work so that he could return to it fresh on Mondays. Dirac's
supervisor, Fowler, introduced him to atomic theory and Dirac spent many
of his days reading Arnold Sommerfeld's book on atomic theory.
Sommerfeld improved on the Bohr model of the atom by using Einstein's
relativistic mechanics instead of Newtonian mechanics to describe the
orbit of electrons around the nucleus, and Sommerfeld was able to obtain
more accurate calculations of the energy levels of the atom compared to
Bohr's original model. Dirac, being interested in relativity, must have
been influenced greatly by this work, since years later he would
merge relativity and quantum mechanics.

@@

What was strange to Heisenberg was that the order in which matrices are
multiplied matters, $a$ times $b$ could be different from $b$ times $a$.
This $ab - ba$ caught Dirac's attention, as Dirac was familiar with this
noncommuting multiplication from his mathematics training, but did not
understand its physical significance or implications. Always interested
in relativity, Dirac also tried to create a relativistic version of
Heisenberg's theory but had no success and soon gave up.

@@

>>>>

\begin{document}
\begin{center}
\textbf{P. A. M. Dirac and his Search for Mathematical Beauty in Nature}
\end{center}

@@

Dirac began studying engineering at the University of Bristol when he
was 16. During his time at the university, he read in a newspaper about
the verification of Einstein's general theory of relativity, which had a
huge influence on him and sparked his interest in physics. Eager to find
out more about Einstein's theory, Dirac was disappointed when he
realized that none of his teachers knew much about relativity either.
Fortunately for Dirac, the philosopher Charlie Broad, who had studied
the Einstein's theory, was appointed as the Professor of Philosphy at
Bristol in Dirac's final year, and gave lectures on relativity. From
these lectures and from reading Arthur Eddington's book on relativity,
Dirac was able to learn both the special and general theory of
relativity.

While Einstein's theories sparked Dirac's interest in physics, Dirac's
education in engineering helped him develop an appreciation for
approximate theories, because engineers had to constantly make
approximations. Dirac realized that Newtonian mechanics can be viewed as
an approximation to Einstein's more refined relativistic mechanics, and
perhaps Einstein's theory is also an approximation to reality. Dirac
developed the belief that the purpose of science is to develop ever more
accurate approximations, and that an exact theory might never be found.
Dirac thought that ``even approximate theories can have mathematical
beauty'' \cite[pg. 45]{strangest_man} as a result of his engineering
training, and this point of view of science as approximations has
influenced many other physicists today.

@@

Dirac was a hard-working student at Cambridge. He worked almost all day
for six days every week, reading books, journals, and reviewing his
lecture notes, only taking time off on Sundays. On Sundays, he would
take long walks that lasted the whole day, and tried not to think about
his work so that he could return to it fresh on Mondays. Dirac's
supervisor, Fowler, introduced him to atomic theory and Dirac spent many
of his days reading Arnold Sommerfeld's book on atomic theory.
Sommerfeld improved on the Bohr model of the atom by using Einstein's
relativistic mechanics instead of Newtonian mechanics to describe the
orbit of electrons around the nucleus, and Sommerfeld was able to obtain
more accurate calculations of the energy levels of the atom compared to
Bohr's original model. Dirac, being interested in relativity, must have
been influenced greatly by this work, since years later he would
merge special relativity and quantum mechanics.

@@

Heisenberg's theory was strange to physicists because the order in which
matrices are multiplied matters, $a$ times $b$ could be different from
$b$ times $a$, so that $ab - ba$ is not zero..  This $ab - ba$ caught
Dirac's attention, as Dirac was familiar with this noncommuting
multiplication from his mathematics training, but did not understand its
physical significance or implications. Always interested in relativity,
Dirac also tried to create a relativistic version of Heisenberg's theory
but had no success and soon gave up.

@@

When Dirac first heard of Schrodinger's wave mechanics, he did not take
it very seriously, as he already had a working quantum mechanics from
Heisenberg and his own work. However, Heisenberg told Dirac to pay more
attention to Schrodinger's version of quantum mechanics. From reading
Schrodinger and other people's work, he realized that he could have
gotten Schrodinger's equation from Heisenberg's theory and his own
developments with Poisson brackets.

Working with Schrodinger's wavefunctions, Dirac thought about how to
represent two or more identical particles using wavefunctions. Dirac
noticed that if two identical particles exchanged places, no experiment
could possibly be done to tell that they have swapped. Following
Heisenberg's influence of only talking about things that could be
measured, Dirac realized that quantum theory should not predict a
measurable change when two identical particles are swapped. Dirac worked
out that the wavefunction must either stay the same or be multiplied by
$-1$ when two identical particles swap in order for quantum theory to be
consistent with this fact. Dirac found that fermions' wavefunctions
would be multiplied by $-1$ while bosons' wavefunctions would stay the
same. Fermions, an example being the electron, are particles that have
the property that no two fermions can occupy the same state (the Pauli
exclusion principle), and in some sense have a tendency to avoid each
other.  Bosons, an example being the photon, are particles that can have
infinitely many of them occupying the same state, and in some sense tend
to clump.

Continuing to play with the mathematics of Schrodinger's equation, Dirac
wondered what would happen if he made the wavefunction itself one of
Heisenberg's noncommuting quantities. From this line of thought, Dirac
worked out a novel mathematical way to represent fermions and bosons,
and this method is now called second quantization. Using this method, he
was able to explain Planck's blackbody radiation formula using the new
quantum mechanics by identifying bosons with oscillators. Second
quantization has become one of Dirac's greatest contributions to
physics, and it came from one of Dirac's unusual approaches to physics
of playing with equations, looking for beautiful mathematics, and seeing
what they predicted, as opposed to having a clear goal in mind and
trying to develop a method for it.

Dirac's successes in working with quantum mechanics must have given him
confidence to return to his old passion, relativity. Dirac now tackled
the problem of merging special relativity with quantum mechanics. Dirac
again resorted to playing with mathematical quantities but this time had
a clear goal which helped him limit the possibilities. He specified the
characteristics the equation must have, and used those to guide him. The
equation must agree with Einstein's special relativity, and also must
agree with his version of quantum mechanics. Dirac also believed that
the equation would be simple and beautiful. Through trial and error and
months of hard work, Dirac eventually guessed the correct equation, now
called the Dirac equation. Unlike Schrodinger's equation which only
includes one wavefunction, Dirac's equation featured four interconnected
wavefunctions. Additionally, Dirac's equation explained the existence of
spin and magnetic field of the electron as a result of the merging of
special relativity and quantum mechanics, a great triumph for
theoretical physics.

All was not well, however. Dirac's equation also predicted the existence
of negative energy electron states, something that seemed absurd and was
never observed in experiment before. This troubled Dirac greatly. He
eventually postulated that all the negative energy states were filled
somehow with undetected negative energy electrons. A vacancy in these
negative energy states would appear as a positively charged hole, and
only this hole could be detected in experiment. In 1932 at Caltech, Carl
Anderson photographed the particle track of Dirac's hole, now called the
positron. In a way, Dirac for the first time had predicted the existence
of a new particle of nature from guessing an equation that would conform
with previously known laws (quantum mechanics and relativity). This
style of theoretical physics, of guessing the mathematics based on
constraints, continued to help physicists predict the existence of more
particles, with the Higgs boson being the most recently discovered one,
but having been predicted by theory many years earlier.

Dirac's successes in producing physical theories from playing with
mathematics led him to believe that the ``most powerful method of
advance...\ is to employ all the resources of pure mathematics in
attempts to perfect and generalise the mathematical formalism that forms
the existing basis of theoretical physics, and \emph{after} each success
in this direction, to try to interpret the new mathematical features in
terms of physical entities'' \cite{monopoles}.

Following his own advice, Dirac generalized the mathematical form of
Schrodinger's wavefunctions and reinterpreted his new wavefunction as
describing particles in a magnetic field. Investigating the properties
of these generalized wavefunctions further, Dirac discovered that if
magnetic monopoles existed, it would explain the quantization of
electric charge. Physicists have only found free particles or ions with
an electric charge being an integer multiple of the charge of an
electron, but now Dirac has a theoretical explanation for this fact.
Unfortunately, his explanation rests on the existence of magnetic
monopoles, and to this day no experiment has found one yet.
